---
title: "Customer Status Prediction"
author: "Bartu Paçal, Begüm Başovalı"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: darkly
    highlight: tango
---


# Customer Status Prediction: Training Walkthrough

## 0) Setup
- Clear the workspace to avoid stale objects, fix a seed for reproducibility, and define project/data paths.
- Load the packages used across EDA, preprocessing, and modeling.

```{r setup, message=FALSE, warning=FALSE}
rm(list = ls())
set.seed(42)

PROJECT_DIR <- "C:/Users/Bartu/Desktop/Customer_Status_Prediction"
DATA_DIR <- file.path(PROJECT_DIR, "Train Datasets")

library(dplyr)
library(tidyr)
library(caret)
library(corrplot)
library(ggplot2)
library(randomForest)
library(nnet)
library(gbm)
library(psych)
library(scales)
library(pROC)
library(tidytext)
```

cat("=========================================================\n")
cat("CUSTOMER STATUS PREDICTION - TRAIN PHASE\n")
cat("=========================================================\n")

## 1) Read Data
- Set up a helper to load CSVs from `DATA_DIR` and keep column types as character/numeric (no automatic factors).
- List all train files in one place so it’s easy to add/remove sources.
- Load into a named list, then unpack into the global environment for downstream steps.

```{r read-data, message=FALSE, warning=FALSE}
cat("\nReading data files...\n")

read_data_file <- function(file_name) {
  read.csv(file.path(DATA_DIR, file_name), stringsAsFactors = FALSE)
}

data_files <- list(
  status = "customer_status_level_train.csv",
  demographics = "customer_demographics_train.csv",
  revenue = "customer_monthly_recurring_revenue_train.csv",
  region = "customer_region_and_industry_train.csv",
  history = "customer_revenue_history_train.csv",
  satisfaction = "customer_satisfaction_scores_train.csv",
  engagement = "newsletter_engagement_train.csv",
  bug = "product_bug_reports_train.csv",
  ticket = "support_ticket_activity_train.csv"
)

data_list <- lapply(data_files, read_data_file)
names(data_list) <- c(
  "Customer_Status", "Customer_Demographics", "Customer_Revenue",
  "Customer_Region_Industry", "Customer_Revenue_History",
  "Customer_Satisfaction", "Customer_Engagement",
  "Customer_Bug_Reports", "Customer_Ticket_Activity"
)

list2env(data_list, envir = .GlobalEnv)
cat("Data files loaded.\n")
```

## 2) Standardize ID Column Names
- Align key columns to a single `Customer.ID` naming so joins/merges stay consistent across tables.
- Rename once up front to avoid scattered fixes later.

```{r id-standardize, message=FALSE, warning=FALSE}
Customer_Demographics <- Customer_Demographics %>%
  rename(Customer.ID = CUS.ID)
```

## 3) Duplicate Record Check
- Build a quick audit table: row/column counts, NA totals, and unique customer counts per table.
- Inspect per-column types, NA rates, and cardinality to spot bad or sparse fields.
- Review target balance, then count duplicate IDs per dataset.
```{r , message=FALSE, warning=FALSE}

cat("\nDuplicate record check...\n")
cat("\nStarting data set preliminary examination...\n")

data_summary <- data.frame(
  Dataset = names(data_list),
  Rows = sapply(data_list, nrow),
  Cols = sapply(data_list, ncol),
  Total_NA = sapply(data_list, function(x) sum(is.na(x))),
  Unique_Customers = sapply(data_list, function(x) {
    id_col <- grep("ID", names(x), value = TRUE, ignore.case = TRUE)[1]
    if (!is.na(id_col)) length(unique(x[[id_col]])) else NA
  })
)

print(data_summary)

cat("\nDetailed column and type analysis:\n")

check_details <- function(df_name, df) {
  cat(paste0("\n--- Table: ", df_name, " ---\n"))
  details <- data.frame(
    Column = names(df),
    Type = sapply(df, class),
    NA_Count = sapply(df, function(x) sum(is.na(x))),
    NA_Perc = sapply(df, function(x) round(mean(is.na(x)) * 100, 2)),
    Unique_Values = sapply(df, function(x) length(unique(x)))
  )
  print(details)
}

invisible(lapply(names(data_list), function(n) check_details(n, data_list[[n]])))

cat("\nTarget variable distribution:\n")
status_dist <- table(Customer_Status$Status)
print(status_dist)
print(prop.table(status_dist))

ggplot(Customer_Status, aes(x = Status, fill = Status)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "Customer Status Distribution", x = "Status", y = "Count")

dup_counts <- sapply(data_list, function(x) {
  id_col <- grep("ID", names(x), value = TRUE, ignore.case = TRUE)[1]
  if (!is.na(id_col)) sum(duplicated(x[[id_col]])) else 0
})

dup_summary <- data.frame(
  Dataset = names(dup_counts),
  Duplicate_IDs = as.numeric(dup_counts)
)

cat("\nDuplicate ID counts:\n")
print(dup_summary)
```

## 4) Aggregation - Customer-based Summarization
- Müşteri bazında özetler: kategorikler için mod, sayımlar için toplam, sürekli skorlar için ortalama.
- Son memnuniyet anketinin tarih/çeyrek bilgisini de sakla.
- Tekrarlı satır yoksa ilgili tabloları olduğu gibi bırak.

```{r agg-customer, message=FALSE, warning=FALSE}
cat("\nData aggregation operations...\n")

get_mode_chr <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_character_)
  tb <- sort(table(x), decreasing = TRUE)
  names(tb)[1]
}

Customer_Bug_Reports_Agg <- Customer_Bug_Reports %>%
  group_by(Customer.ID) %>%
  summarise(
    Total_Bug_Count = sum(Product.Bug.Task.Count, na.rm = TRUE),
    .groups = "drop"
  )

names(Customer_Satisfaction) <- c(
  "Customer.ID", "Year", "Quarter", "Survey_Date", "Response_Date",
  "NPS", "Value_Score", "Usage_Freq", "Quality_Score",
  "Usability_Score", "Reporting_Score"
)

Customer_Satisfaction <- Customer_Satisfaction %>%
  mutate(
    Survey_Date = as.Date(Survey_Date),
    Response_Date = as.Date(Response_Date)
  )

last_rows <- Customer_Satisfaction %>%
  mutate(last_date = coalesce(Survey_Date, Response_Date)) %>%
  arrange(Customer.ID, last_date) %>%
  group_by(Customer.ID) %>%
  slice_tail(n = 1) %>%
  ungroup() %>%
  select(Customer.ID, Year, Quarter, Survey_Date, Response_Date)

Customer_Satisfaction_Agg <- Customer_Satisfaction %>%
  group_by(Customer.ID) %>%
  summarise(
    NPS_mean = mean(NPS, na.rm = TRUE),
    Value_mean = mean(Value_Score, na.rm = TRUE),
    Quality_mean = mean(Quality_Score, na.rm = TRUE),
    Usability_mean = mean(Usability_Score, na.rm = TRUE),
    Reporting_score = get_mode_chr(Reporting_Score),
    Usage_Freq = get_mode_chr(Usage_Freq),
    Survey_count = n(),
    .groups = "drop"
  ) %>%
  left_join(
    last_rows %>%
      rename(
        Year_Last = Year,
        Quarter_Last = Quarter,
        Last_Survey_Date = Survey_Date,
        Last_Response_Date = Response_Date
      ),
    by = "Customer.ID"
  )

Customer_Ticket_Activity_Agg <- if (sum(duplicated(Customer_Ticket_Activity$Customer.ID)) > 0) {
  Customer_Ticket_Activity %>%
    group_by(Customer.ID) %>%
    summarise(
      Help.Ticket.Count = sum(Help.Ticket.Count, na.rm = TRUE),
      Help.Ticket.Lead.Time..hours. = mean(Help.Ticket.Lead.Time..hours., na.rm = TRUE),
      .groups = "drop"
    )
} else {
  Customer_Ticket_Activity
}

Customer_Engagement_Agg <- if (sum(duplicated(Customer_Engagement$Customer.ID)) > 0) {
  Customer_Engagement %>%
    group_by(Customer.ID) %>%
    summarise(
      Company.Newsletter.Interaction.Count = sum(Company.Newsletter.Interaction.Count, na.rm = TRUE),
      .groups = "drop"
    )
} else {
  Customer_Engagement
}

cat("Aggregation operations completed.\n")
```
## 5) Merge All Data Sets (master_train)
- Birleştirme anahtarı: tüm tablolar Customer.ID ile left_join() yapılıyor; Customer_Status’taki tüm müşteriler korunur.
- Doğrulama: satır sayısı ve Customer.ID tekilliği kontrol edilir.
- Hızlı numerik özet: yalnızca sayısal kolonlar için psych::describe().
```{r , message=FALSE, warning=FALSE}
cat("\nData merging...\n")

master_train <- Customer_Status %>%
  left_join(Customer_Demographics, by = "Customer.ID") %>%
  left_join(Customer_Revenue, by = "Customer.ID") %>%
  left_join(Customer_Region_Industry, by = "Customer.ID") %>%
  left_join(Customer_Revenue_History, by = "Customer.ID") %>%
  left_join(Customer_Satisfaction_Agg, by = "Customer.ID") %>%
  left_join(Customer_Engagement_Agg, by = "Customer.ID") %>%
  left_join(Customer_Bug_Reports_Agg, by = "Customer.ID") %>%
  left_join(Customer_Ticket_Activity_Agg, by = "Customer.ID")

stopifnot(nrow(master_train) == nrow(Customer_Status))
stopifnot(sum(duplicated(master_train$Customer.ID)) == 0)

cat("Data merged. Total customers:", nrow(master_train), "\n")

cat("\nNumeric variable summary:\n")
num_cols <- master_train %>% select(where(is.numeric))
psych::describe(num_cols)
```

## 6) Basic Cleaning (Before Correlation Analysis)
- Para alanları: MRR, Total.Revenue içindeki $, karakterleri temizlenip numeriğe çevrilir.
- Kategorikler: modelleme için faktöre çevrilir (Status, Region, vb.).
- Tarihler: Last_Survey_Date, Last_Response_Date Date tipine dönüştürülür.

```{r , message=FALSE, warning=FALSE}
cat("\nBasic data cleaning...\n")

currency_columns <- c("MRR", "Total.Revenue")
for (col in currency_columns) {
  if (col %in% colnames(master_train)) {
    master_train[[col]] <- as.numeric(gsub("[\\$,]", "", as.character(master_train[[col]])))
  }
}

factor_columns <- c("Status", "Region", "Customer.Level", "Vertical", "Subvertical", 
                    "Usage_Freq", "Reporting_score")
for (col in factor_columns) {
  if (col %in% colnames(master_train)) {
    master_train[[col]] <- as.factor(master_train[[col]])
  }
}

date_columns <- c("Last_Survey_Date", "Last_Response_Date")
for (col in date_columns) {
  if (col %in% colnames(master_train)) {
    master_train[[col]] <- as.Date(master_train[[col]])
  }
}

cat("Basic cleaning completed.\n")
```

7) Correlation Analysis (Before Feature Engineering)
- Numeric selection: choose numeric features and drop Year_Last, Quarter_Last (categorical-like numerics).
- Constant removal: discard zero-variance columns to avoid degenerate correlations.
- Matrix + heatmap: compute Pearson correlations (complete vs. pairwise depending on missingness) and save a heatmap.
- High correlations: list pairs with |r| > 0.8, save CSV, and export top-3 scatter plots.
- Status link: convert Status to numeric; compute per-feature correlation and save a ranked bar chart.

```{r , message=FALSE, warning=FALSE}
cat("\n", strrep("=", 60), "\n")
cat("CORRELATION ANALYSIS - Before Feature Engineering\n")
cat(strrep("=", 60), "\n")

numeric_vars_raw <- master_train %>%
  select(where(is.numeric)) %>%
  select(-Year_Last, -Quarter_Last)

zero_sd_vars <- names(which(apply(numeric_vars_raw, 2, function(x) sd(x, na.rm = TRUE)) == 0))
if (length(zero_sd_vars) > 0) {
  cat("Removing constant value variables:", paste(zero_sd_vars, collapse = ", "), "\n")
  numeric_vars_raw <- numeric_vars_raw %>% select(-all_of(zero_sd_vars))
}

cat("Number of numeric variables analyzed:", ncol(numeric_vars_raw), "\n")

complete_cases <- complete.cases(numeric_vars_raw)
if (sum(complete_cases) < 10) {
  cat("Too much missing data, using pairwise correlation\n")
  cor_matrix_raw <- cor(numeric_vars_raw, use = "pairwise.complete.obs", method = "pearson")
} else {
  cor_matrix_raw <- cor(numeric_vars_raw[complete_cases, ], use = "complete.obs", method = "pearson")
}

cat("Correlation matrix dimensions:", dim(cor_matrix_raw), "\n")

cat("\nCreating raw correlation heatmap...\n")
png(file.path(PROJECT_DIR, "correlation_raw_before_fe.png"), width = 1200, height = 1000)
corrplot(cor_matrix_raw, 
         method = "color",
         type = "upper",
         tl.col = "black",
         tl.cex = 0.7,
         number.cex = 0.6,
         title = "RAW CORRELATION MATRIX (Before Feature Engineering)",
         mar = c(0, 0, 2, 0))
dev.off()
cat("Raw correlation heatmap saved: correlation_raw_before_fe.png\n")

cor_matrix_raw_no_na <- cor_matrix_raw
cor_matrix_raw_no_na[is.na(cor_matrix_raw_no_na)] <- 0

high_corr_raw <- which(abs(cor_matrix_raw_no_na) > 0.8 & 
                         upper.tri(cor_matrix_raw_no_na) & 
                         cor_matrix_raw_no_na != 1, arr.ind = TRUE)

if (length(high_corr_raw) > 0 && nrow(high_corr_raw) > 0) {
  cat("\nHIGHLY CORRELATED VARIABLES (|r| > 0.8):\n")
  high_corr_df_raw <- data.frame(
    Variable1 = rownames(cor_matrix_raw_no_na)[high_corr_raw[, 1]],
    Variable2 = colnames(cor_matrix_raw_no_na)[high_corr_raw[, 2]],
    Correlation = cor_matrix_raw_no_na[high_corr_raw]
  )
  high_corr_df_raw <- high_corr_df_raw[order(-abs(high_corr_df_raw$Correlation)), ]
  print(high_corr_df_raw)
  
  write.csv(high_corr_df_raw, file.path(PROJECT_DIR, "high_correlation_raw.csv"), row.names = FALSE)
  cat("\nRaw correlation analysis saved: high_correlation_raw.csv\n")
  
  cat("\nHIGH CORRELATION ANALYSIS:\n")
  top_pairs <- head(high_corr_df_raw, 3)
  
  for (i in 1:min(3, nrow(top_pairs))) {
    var1 <- top_pairs$Variable1[i]
    var2 <- top_pairs$Variable2[i]
    corr <- top_pairs$Correlation[i]
    
    plot_data <- master_train[complete.cases(master_train[, c(var1, var2)]), c(var1, var2)]
    
    if (nrow(plot_data) > 10) {
      p <- ggplot(plot_data, aes_string(x = var1, y = var2)) +
        geom_point(alpha = 0.5, color = "steelblue", size = 2) +
        geom_smooth(method = "lm", color = "red", se = FALSE) +
        labs(
          title = paste("High Correlation: ", round(corr, 3)),
          subtitle = paste(var1, "vs", var2),
          x = var1,
          y = var2
        ) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, face = "bold"),
              plot.subtitle = element_text(hjust = 0.5))
      
      print(p)
      
      filename <- paste0("high_corr_", gsub("[^A-Za-z0-9]", "_", var1), 
                         "_vs_", gsub("[^A-Za-z0-9]", "_", var2), ".png")
      ggsave(file.path(PROJECT_DIR, filename), plot = p, width = 8, height = 6)
    }
  }
  cat("High correlation scatter plots saved\n")
} else {
  cat("\nNo highly correlated variables in raw data (|r| > 0.8)\n")
}

cat("\nCORRELATION ANALYSIS WITH STATUS:\n")

status_numeric <- as.numeric(master_train$Status)

status_correlations_raw <- data.frame(
  Variable = colnames(numeric_vars_raw),
  Correlation_with_Status = NA_real_,
  Abs_Correlation = NA_real_
)

for (i in 1:ncol(numeric_vars_raw)) {
  var_name <- colnames(numeric_vars_raw)[i]
  var_data <- numeric_vars_raw[[i]]
  
  valid_indices <- complete.cases(data.frame(var_data, status_numeric))
  
  if (sum(valid_indices) > 10 && sd(var_data[valid_indices], na.rm = TRUE) > 0) {
    cor_result <- cor(var_data[valid_indices], status_numeric[valid_indices], use = "complete.obs")
    status_correlations_raw$Correlation_with_Status[i] <- cor_result
    status_correlations_raw$Abs_Correlation[i] <- abs(cor_result)
  }
}

status_correlations_raw <- status_correlations_raw %>%
  filter(!is.na(Correlation_with_Status)) %>%
  arrange(-Abs_Correlation)

cat("\nVARIABLES WITH HIGHEST CORRELATION WITH STATUS:\n")
print(head(status_correlations_raw, 10))

n_vars_to_plot <- min(15, nrow(status_correlations_raw))
if (n_vars_to_plot > 0) {
  plot_data <- head(status_correlations_raw, n_vars_to_plot)
  
  p_status <- ggplot(plot_data, 
                     aes(x = reorder(Variable, Abs_Correlation), y = Correlation_with_Status)) +
    geom_bar(stat = "identity", 
             fill = ifelse(plot_data$Correlation_with_Status > 0, "steelblue", "coral"), 
             alpha = 0.8) +
    geom_text(aes(label = sprintf("%.3f", Correlation_with_Status)), 
              hjust = ifelse(plot_data$Correlation_with_Status > 0, -0.1, 1.1),
              size = 3) +
    coord_flip() +
    labs(
      title = "Correlation with Status",
      x = "Variable",
      y = "Correlation Coefficient"
    ) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
  
  print(p_status)
  ggsave(file.path(PROJECT_DIR, "status_correlation.png"), plot = p_status, width = 10, height = 8)
}
```

## 8) Feature Engineering (Based on Correlation Analysis)
- Findings summary: note highly correlated pairs and variables most related to Status.
- Missing values: add flags and impute (MRR, Total.Revenue, Customer.Age..Months.).
- Survey scores: has_survey flag and median imputation for NPS_mean, Value_mean, Quality_mean, Usability_mean.
- Categorical NA: fill Reporting_score, Usage_Freq with "None".
- Response delay: add no_response_flag and response_delay_days.
- Engagement/bug/ticket: presence flags and set missing counts to 0.
- New features: Satisfaction Composite, Support Burden, Revenue per Month, Bug Frequency, Engagement Ratio, MRR Ratio.

```{r , message=FALSE, warning=FALSE}
cat("\n", strrep("=", 60), "\n")
cat("FEATURE ENGINEERING - Based on Correlation Analysis\n")
cat(strrep("=", 60), "\n")

cat("\nCORRELATION ANALYSIS FINDINGS:\n")
if (exists("high_corr_df_raw") && nrow(high_corr_df_raw) > 0) {
  cat("1. Highly correlated variable pairs:", nrow(high_corr_df_raw), "pairs\n")
  cat("   Example:", paste(head(high_corr_df_raw$Variable1, 2), 
                           head(high_corr_df_raw$Variable2, 2), sep = " vs ", collapse = ", "), "\n")
} else {
  cat("1. No highly correlated variable pairs\n")
}
cat("2. Variables most related to Status identified\n")
cat("3. Feature engineering will use these insights\n")

cat("\nMISSING VALUE PROCESSING...\n")

# Flags + imputation for core numeric fields
master_train$MRR_na_flag <- as.integer(is.na(master_train$MRR))
master_train$MRR[is.na(master_train$MRR)] <- 0

master_train$Total.Revenue_na_flag <- as.integer(is.na(master_train$Total.Revenue))
master_train$Total.Revenue[is.na(master_train$Total.Revenue)] <- 0

if ("Customer.Age..Months." %in% colnames(master_train)) {
  master_train$Customer.Age..Months.[is.na(master_train$Customer.Age..Months.)] <- 
    median(master_train$Customer.Age..Months., na.rm = TRUE)
}

# Survey presence + targeted imputation
master_train$has_survey <- ifelse(is.na(master_train$NPS_mean), 0, 1)

survey_vars <- c("NPS_mean", "Value_mean", "Quality_mean", "Usability_mean")
for (var in survey_vars) {
  if (var %in% colnames(master_train)) {
    master_train[[var]][is.na(master_train[[var]])] <- median(master_train[[var]], na.rm = TRUE)
  }
}

# Categorical NA -> "None"
master_train <- master_train %>%
  mutate(
    Reporting_score = replace_na(as.character(Reporting_score), "None"),
    Usage_Freq = replace_na(as.character(Usage_Freq), "None")
  )

# Survey_count NA -> 0
master_train$Survey_count[is.na(master_train$Survey_count)] <- 0

# Response delay features
master_train$no_response_flag <- ifelse(
  is.na(master_train$Last_Response_Date) | is.na(master_train$Last_Survey_Date), 1, 0
)

master_train$response_delay_days <- as.numeric(master_train$Last_Response_Date - master_train$Last_Survey_Date)
master_train$response_delay_days[is.na(master_train$response_delay_days)] <- 0
master_train$response_delay_days[master_train$response_delay_days < 0] <- 0

# Engagement / bug / ticket flags + counts
master_train$has_newsletter_interaction <- ifelse(is.na(master_train$Company.Newsletter.Interaction.Count), 0, 1)
master_train$Company.Newsletter.Interaction.Count[is.na(master_train$Company.Newsletter.Interaction.Count)] <- 0

master_train$has_bug_report <- ifelse(is.na(master_train$Total_Bug_Count), 0, 1)
master_train$Total_Bug_Count[is.na(master_train$Total_Bug_Count)] <- 0

master_train$has_ticket <- ifelse(is.na(master_train$Help.Ticket.Count), 0, 1)
master_train$Help.Ticket.Count[is.na(master_train$Help.Ticket.Count)] <- 0

master_train$Help.Ticket.Lead.Time..hours.[is.na(master_train$Help.Ticket.Lead.Time..hours.)] <- 0

cat("\nCreating new features...\n")

# Satisfaction Composite
master_train$Satisfaction_Composite <- rowMeans(
  master_train[, c("NPS_mean", "Value_mean", "Quality_mean", "Usability_mean")],
  na.rm = TRUE
)

# Support Burden
master_train$Support_Burden <- master_train$Help.Ticket.Count * 
  ifelse(master_train$Help.Ticket.Lead.Time..hours. > 0, 
         master_train$Help.Ticket.Lead.Time..hours., 1)

# Revenue per month
master_train$Revenue_per_Month <- ifelse(
  master_train$Customer.Age..Months. > 0,
  master_train$Total.Revenue / master_train$Customer.Age..Months.,
  0
)

# Bug frequency
master_train$Bug_Frequency <- ifelse(
  master_train$Customer.Age..Months. > 0,
  master_train$Total_Bug_Count / master_train$Customer.Age..Months.,
  0
)

# Engagement ratio
master_train$Engagement_Ratio <- ifelse(
  master_train$Customer.Age..Months. > 0,
  master_train$Company.Newsletter.Interaction.Count / master_train$Customer.Age..Months.,
  0
)

# MRR to revenue ratio
master_train$MRR_Ratio <- ifelse(
  master_train$Total.Revenue > 0,
  master_train$MRR / master_train$Total.Revenue,
  0
)

cat("6 new features created.\n")
```

## 9) Post-Correlation Analysis Check
- Numeric selection: re-select numeric features and drop Year_Last, Quarter_Last.
- Constant removal: remove zero-variance columns.
- Final matrix + heatmap: compute Pearson correlation and save a final heatmap.
- High correlations: list |r| > 0.8 pairs, sort, and save to CSV.
- Status link: compute correlations of final features with Status; visualize new features.

```{r , message=FALSE, warning=FALSE}
cat("\n", strrep("=", 60), "\n")
cat("CORRELATION ANALYSIS - After Feature Engineering\n")
cat(strrep("=", 60), "\n")

numeric_vars_final <- master_train %>%
  select(where(is.numeric)) %>%
  select(-Year_Last, -Quarter_Last)

zero_sd_vars_final <- names(which(apply(numeric_vars_final, 2, function(x) sd(x, na.rm = TRUE)) == 0))
if (length(zero_sd_vars_final) > 0) {
  cat("Removing constant value variables:", paste(zero_sd_vars_final, collapse = ", "), "\n")
  numeric_vars_final <- numeric_vars_final %>% select(-all_of(zero_sd_vars_final))
}

cat("Number of numeric variables after feature engineering:", ncol(numeric_vars_final), "\n")

complete_cases_final <- complete.cases(numeric_vars_final)
if (sum(complete_cases_final) < 10) {
  cor_matrix_final <- cor(numeric_vars_final, use = "pairwise.complete.obs", method = "pearson")
} else {
  cor_matrix_final <- cor(numeric_vars_final[complete_cases_final, ], use = "complete.obs", method = "pearson")
}

png(file.path(PROJECT_DIR, "correlation_final_after_fe.png"), width = 1200, height = 1000)
corrplot(cor_matrix_final, 
         method = "color",
         type = "upper",
         tl.col = "black",
         tl.cex = 0.6,
         number.cex = 0.5,
         title = "FINAL CORRELATION MATRIX (After Feature Engineering)",
         mar = c(0, 0, 2, 0))
dev.off()
cat("Final correlation heatmap saved: correlation_final_after_fe.png\n")

cor_matrix_final_no_na <- cor_matrix_final
cor_matrix_final_no_na[is.na(cor_matrix_final_no_na)] <- 0

high_corr_final <- which(abs(cor_matrix_final_no_na) > 0.8 & 
                           upper.tri(cor_matrix_final_no_na) & 
                           cor_matrix_final_no_na != 1, arr.ind = TRUE)

if (length(high_corr_final) > 0 && nrow(high_corr_final) > 0) {
  cat("\nHIGHLY CORRELATED VARIABLES AFTER FEATURE ENGINEERING (|r| > 0.8):\n")
  high_corr_df_final <- data.frame(
    Variable1 = rownames(cor_matrix_final_no_na)[high_corr_final[, 1]],
    Variable2 = colnames(cor_matrix_final_no_na)[high_corr_final[, 2]],
    Correlation = cor_matrix_final_no_na[high_corr_final]
  )
  high_corr_df_final <- high_corr_df_final[order(-abs(high_corr_df_final$Correlation)), ]
  print(high_corr_df_final)
  
  write.csv(high_corr_df_final, file.path(PROJECT_DIR, "high_correlation_final.csv"), row.names = FALSE)
  cat("\nFinal correlation analysis saved: high_correlation_final.csv\n")
} else {
  cat("\nNo highly correlated variables after feature engineering (|r| > 0.8)\n")
}

status_correlations_final <- data.frame(
  Variable = colnames(numeric_vars_final),
  Correlation_with_Status = NA_real_,
  Abs_Correlation = NA_real_
)

for (i in 1:ncol(numeric_vars_final)) {
  var_name <- colnames(numeric_vars_final)[i]
  var_data <- numeric_vars_final[[i]]
  
  valid_indices <- complete.cases(data.frame(var_data, status_numeric))
  
  if (sum(valid_indices) > 10 && sd(var_data[valid_indices], na.rm = TRUE) > 0) {
    cor_result <- cor(var_data[valid_indices], status_numeric[valid_indices], 
                      use = "complete.obs")
    status_correlations_final$Correlation_with_Status[i] <- cor_result
    status_correlations_final$Abs_Correlation[i] <- abs(cor_result)
  }
}

status_correlations_final <- status_correlations_final %>%
  filter(!is.na(Correlation_with_Status)) %>%
  arrange(-Abs_Correlation)

cat("\nNEW VARIABLES WITH HIGHEST CORRELATION WITH STATUS:\n")
print(head(status_correlations_final, 10))

new_features <- c("Satisfaction_Composite", "Support_Burden", "Revenue_per_Month", 
                  "Bug_Frequency", "Engagement_Ratio", "MRR_Ratio")
new_features_data <- status_correlations_final %>%
  filter(Variable %in% new_features)

if (nrow(new_features_data) > 0) {
  p_new_features <- ggplot(new_features_data, 
                           aes(x = reorder(Variable, Abs_Correlation), y = Correlation_with_Status)) +
    geom_bar(stat = "identity", fill = "lightgreen", alpha = 0.8) +
    geom_text(aes(label = sprintf("%.3f", Correlation_with_Status)), 
              hjust = ifelse(new_features_data$Correlation_with_Status > 0, -0.1, 1.1),
              size = 3.5) +
    coord_flip() +
    labs(
      title = "New Features Correlation with Status",
      x = "New Feature",
      y = "Correlation Coefficient"
    ) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
  
  print(p_new_features)
  ggsave(file.path(PROJECT_DIR, "new_features_correlation.png"), 
         plot = p_new_features, width = 10, height = 6)
}
```
## 10) Variable Selection and Model Preparation
- Selection criteria:
 - Variables most correlated with Status
 - Remove one from highly correlated pairs (multicollinearity control)
- Include newly engineered features
-Exclusions:
 - IDs and dates (Customer.ID, Last_Survey_Date, Last_Response_Date)
 - Original variables replaced by new composites (e.g., NPS_mean, Help.Ticket.Count)
 
```{r , message=FALSE, warning=FALSE}
 cat("\n", strrep("=", 60), "\n")
cat("VARIABLE SELECTION AND MODEL PREPARATION\n")
cat(strrep("=", 60), "\n")

cat("\nVARIABLE SELECTION CRITERIA:\n")
cat("1. Variables with highest correlation with Status\n")
cat("2. Remove one from highly correlated pairs\n")
cat("3. Include newly created features\n")

# Base selection: drop ID and date fields
master_model <- master_train %>%
  select(-Customer.ID, -Last_Survey_Date, -Last_Response_Date)

# Remove original variables that are represented by engineered features
variables_to_remove <- c(
  "NPS_mean", "Value_mean", "Quality_mean", "Usability_mean",
  "Help.Ticket.Count", "Help.Ticket.Lead.Time..hours."
)

cat("\nREMOVING VARIABLES TO PREVENT MULTICOLLINEARITY:\n")
for (var in variables_to_remove) {
  if (var %in% colnames(master_model)) {
    cat("  -", var, "removed\n")
  }
}
master_model <- master_model %>% select(-any_of(variables_to_remove))

# Ensure categorical columns are factors
cat_cols <- c("Region", "Customer.Level", "Vertical", "Subvertical", "Usage_Freq", "Reporting_score")
for (col in cat_cols) {
  if (col %in% colnames(master_model)) {
    master_model[[col]] <- as.factor(as.character(master_model[[col]]))
  }
}

# Drop one of highly correlated numeric variables (cutoff = 0.8)
num_mat <- master_model %>% select(where(is.numeric))
if (ncol(num_mat) > 1) {
  cor_mat <- cor(num_mat, use = "pairwise.complete.obs", method = "pearson")
  cor_mat[is.na(cor_mat)] <- 0
  hc_names <- caret::findCorrelation(cor_mat, cutoff = 0.8, names = TRUE, verbose = FALSE)
  if (length(hc_names) > 0) {
    cat("\nRemoving highly correlated variables (cutoff=0.8): ", paste(hc_names, collapse = ", "), "\n", sep = "")
    master_model <- master_model %>% select(-all_of(hc_names))
  } else {
    cat("\nNo variables removed by high correlation filter.\n")
  }
}

# NA handling: numeric -> median, factor -> mode
cat("\nCHECKING NA VALUES IN FINAL MODEL DATA SET...\n")
na_summary <- sapply(master_model, function(x) sum(is.na(x)))
na_vars <- names(na_summary[na_summary > 0])

if (length(na_vars) > 0) {
  cat("Variables containing NA:", length(na_vars), "\n")
  cat("   Total NA count:", sum(na_summary), "\n")
  
  is_num <- sapply(master_model, is.numeric)
  is_fac <- sapply(master_model, is.factor)
  
  for (var in na_vars) {
    if (isTRUE(is_num[var])) {
      master_model[[var]][is.na(master_model[[var]])] <- median(master_model[[var]], na.rm = TRUE)
      cat("   -", var, ": NA values filled with median (", na_summary[var], " NA)\n", sep = "")
    } else if (isTRUE(is_fac[var])) {
      mode_val <- names(sort(table(master_model[[var]]), decreasing = TRUE))[1]
      master_model[[var]][is.na(master_model[[var]])] <- mode_val
      cat("   -", var, ": NA values filled with mode(", mode_val, ") (", na_summary[var], " NA)\n", sep = "")
    }
  }
}

cat("\nFINAL MODEL DATA SET:\n")
cat("Number of rows:", nrow(master_model), "\n")
cat("Number of columns:", ncol(master_model), "\n")
cat("NA count (after correction):", sum(is.na(master_model)), "\n")

cat("Status distribution:\n")
print(table(master_model$Status))
```

## 11) Post-Cleaning Correlation Analysis (2nd Analysis)
- Numeric selection: use cleaned master_model numerics; drop constant columns.
- Final matrix + heatmap: compute Pearson correlation and save the plot.
- High correlations: list |r| > 0.8 pairs, sort, and save CSV.
- Status link: compute correlations with Status and export a comparison chart.
- Comparison: summarize variable counts and high-correlation reduction vs. earlier phases.

```{r , message=FALSE, warning=FALSE}
cat("\n", strrep("=", 60), "\n")
cat("POST-CLEANING CORRELATION ANALYSIS (2nd Analysis)\n")
cat(strrep("=", 60), "\n")

## 11.1 Select numeric variables after cleaning
numeric_vars_clean <- master_model %>%
  select(where(is.numeric))

# Remove constant value variables
zero_sd_vars_clean <- names(which(apply(numeric_vars_clean, 2, function(x) sd(x, na.rm = TRUE)) == 0))
if (length(zero_sd_vars_clean) > 0) {
  cat("Removing constant value variables:", paste(zero_sd_vars_clean, collapse = ", "), "\n")
  numeric_vars_clean <- numeric_vars_clean %>% select(-all_of(zero_sd_vars_clean))
}

cat("Number of numeric variables after cleaning:", ncol(numeric_vars_clean), "\n")
cat("Variables:", paste(colnames(numeric_vars_clean), collapse = ", "), "\n")

## 11.2 Post-cleaning correlation matrix
complete_cases_clean <- complete.cases(numeric_vars_clean)
if (sum(complete_cases_clean) < 10) {
  cor_matrix_clean <- cor(numeric_vars_clean, use = "pairwise.complete.obs", method = "pearson")
} else {
  cor_matrix_clean <- cor(numeric_vars_clean[complete_cases_clean, ], use = "complete.obs", method = "pearson")
}

cat("Post-cleaning correlation matrix dimensions:", dim(cor_matrix_clean), "\n")

## 11.3 Create post-cleaning heatmap
cat("\nCreating post-cleaning correlation heatmap...\n")
png(file.path(PROJECT_DIR, "correlation_clean_after_removal.png"), width = 1200, height = 1000)
corrplot(cor_matrix_clean, 
         method = "color",
         type = "upper",
         tl.col = "black",
         tl.cex = 0.6,
         number.cex = 0.5,
         title = "POST-CLEANING CORRELATION MATRIX (After Multicollinearity Prevention)",
         mar = c(0, 0, 2, 0))
dev.off()
cat("Post-cleaning correlation heatmap saved: correlation_clean_after_removal.png\n")

## 11.4 Post-cleaning high correlation check
cor_matrix_clean_no_na <- cor_matrix_clean
cor_matrix_clean_no_na[is.na(cor_matrix_clean_no_na)] <- 0

high_corr_clean <- which(abs(cor_matrix_clean_no_na) > 0.8 & 
                           upper.tri(cor_matrix_clean_no_na) & 
                           cor_matrix_clean_no_na != 1, arr.ind = TRUE)

if (length(high_corr_clean) > 0 && nrow(high_corr_clean) > 0) {
  cat("\nHIGHLY CORRELATED VARIABLES AFTER CLEANING (|r| > 0.8):\n")
  high_corr_df_clean <- data.frame(
    Variable1 = rownames(cor_matrix_clean_no_na)[high_corr_clean[, 1]],
    Variable2 = colnames(cor_matrix_clean_no_na)[high_corr_clean[, 2]],
    Correlation = cor_matrix_clean_no_na[high_corr_clean]
  )
  high_corr_df_clean <- high_corr_df_clean[order(-abs(high_corr_df_clean$Correlation)), ]
  print(high_corr_df_clean)
  
  write.csv(high_corr_df_clean, file.path(PROJECT_DIR, "high_correlation_clean.csv"), row.names = FALSE)
  cat("\nPost-cleaning correlation analysis saved: high_correlation_clean.csv\n")
} else {
  cat("\nNo highly correlated variables after cleaning (|r| > 0.8)\n")
}
# Section 11.4'ün sonuna ekleyin
if (!exists("high_corr_df_clean")) {
  high_corr_df_clean <- data.frame(Variable1=character(), Variable2=character(), Correlation=numeric())
}
## 11.5 Post-cleaning Correlation with Status analysis
status_correlations_clean <- data.frame(
  Variable = colnames(numeric_vars_clean),
  Correlation_with_Status = NA_real_,
  Abs_Correlation = NA_real_
)

for (i in 1:ncol(numeric_vars_clean)) {
  var_name <- colnames(numeric_vars_clean)[i]
  var_data <- numeric_vars_clean[[i]]
  
  valid_indices <- complete.cases(data.frame(var_data, status_numeric))
  
  if (sum(valid_indices) > 10 && sd(var_data[valid_indices], na.rm = TRUE) > 0) {
    cor_result <- cor(var_data[valid_indices], status_numeric[valid_indices], 
                      use = "complete.obs")
    status_correlations_clean$Correlation_with_Status[i] <- cor_result
    status_correlations_clean$Abs_Correlation[i] <- abs(cor_result)
  }
}

status_correlations_clean <- status_correlations_clean %>%
  filter(!is.na(Correlation_with_Status)) %>%
  arrange(-Abs_Correlation)

cat("\nVARIABLES WITH HIGHEST CORRELATION WITH STATUS AFTER CLEANING:\n")
print(head(status_correlations_clean, 10))

# Post-cleaning Status correlation graph
n_vars_to_plot_clean <- min(15, nrow(status_correlations_clean))
if (n_vars_to_plot_clean > 0) {
  plot_data_clean <- head(status_correlations_clean, n_vars_to_plot_clean)
  
  p_status_clean <- ggplot(plot_data_clean, 
                           aes(x = reorder(Variable, Abs_Correlation), y = Correlation_with_Status)) +
    geom_bar(stat = "identity", 
             fill = ifelse(plot_data_clean$Correlation_with_Status > 0, "steelblue", "coral"), 
             alpha = 0.8) +
    geom_text(aes(label = sprintf("%.3f", Correlation_with_Status)), 
              hjust = ifelse(plot_data_clean$Correlation_with_Status > 0, -0.1, 1.1),
              size = 3) +
    coord_flip() +
    labs(
      title = "Status Correlation After Cleaning",
      subtitle = "After Multicollinearity Prevention",
      x = "Variable",
      y = "Correlation Coefficient"
    ) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))
  
  print(p_status_clean)
  ggsave(file.path(PROJECT_DIR, "status_correlation_clean.png"), 
         plot = p_status_clean, width = 10, height = 8)
}

## 11.6 Correlation analysis comparison
cat("\nCORRELATION ANALYSIS COMPARISON:\n")
cat("1. Before Feature Engineering: ", ncol(numeric_vars_raw), " variables\n")
cat("2. After Feature Engineering: ", ncol(numeric_vars_final), " variables\n")
cat("3. After Cleaning: ", ncol(numeric_vars_clean), " variables\n")

if (exists("high_corr_df_raw") && exists("high_corr_df_clean")) {
  cat("\nHIGH CORRELATION REDUCTION SUCCESS:\n")
  cat("   Initial high correlation pairs: ", nrow(high_corr_df_raw), "\n")
  cat("   Post-cleaning high correlation pairs: ", nrow(high_corr_df_clean), "\n")
  
  if (nrow(high_corr_df_raw) > 0 && nrow(high_corr_df_clean) > 0) {
    reduction <- round(100 * (1 - nrow(high_corr_df_clean) / nrow(high_corr_df_raw)), 1)
    cat("   High correlation reduction rate: %", reduction, "\n")
  }
}
```
## 12) Train / Validation Split (Stratified)
- Strategy: stratified split by Status with 80/20 ratio using caret::createDataPartition().
- Checks: print set sizes and class distributions to verify stratification.
```{r , message=FALSE, warning=FALSE}
cat("\nTRAIN/VALIDATION SPLIT...\n")

trainIndex <- createDataPartition(master_model$Status, p = 0.8, list = FALSE)
train_data <- master_model[trainIndex, ]
valid_data <- master_model[-trainIndex, ]

cat("Train set size:", nrow(train_data), "\n")
cat("Validation set size:", nrow(valid_data), "\n")
cat("Train distribution:\n")
print(table(train_data$Status))
cat("Validation distribution:\n")
print(table(valid_data$Status))
```

## 13) One‑Hot Encoding (ONLY ON TRAIN)
- Targets: categorical columns Region, Customer.Level, Vertical, Subvertical, Usage_Freq, Reporting_score.
- Method: caret::dummyVars() with fullRank = TRUE to avoid perfect multicollinearity.
- Output: append dummy columns to numeric features for train/validation consistently.

```{r , message=FALSE, warning=FALSE}
cat("\nONE-HOT ENCODING...\n")

# Ensure cat_cols is available
if (!exists("cat_cols")) {
  cat_cols <- c("Region", "Customer.Level", "Vertical", "Subvertical", "Usage_Freq", "Reporting_score")
}

dummies <- dummyVars(
  ~ Region + Customer.Level + Vertical + Subvertical + Usage_Freq + Reporting_score,
  data = train_data,
  fullRank = TRUE
)

train_dummy <- predict(dummies, newdata = train_data) %>% as.data.frame()
valid_dummy <- predict(dummies, newdata = valid_data) %>% as.data.frame()

train_final <- bind_cols(
  train_data %>% select(-all_of(cat_cols)),
  train_dummy
)

valid_final <- bind_cols(
  valid_data %>% select(-all_of(cat_cols)),
  valid_dummy
)

cat("Train size after one-hot encoding:", dim(train_final), "\n")
cat("Validation size after one-hot encoding:", dim(valid_final), "\n")
```

## 14) Model Training
- Control: 5-fold CV, class probabilities, multiClassSummary metrics.
- NA safety: fill remaining NAs (median for numerics) in train_final and mirror to valid_final.
- Models:
- Random Forest (caret::train(method = "rf")) + validation confusion matrix
- Multinomial Logistic (nnet::multinom) + validation confusion matrix
- GBM (caret::train(method = "gbm")) with a small grid + validation confusion matrix

```{r , message=FALSE, warning=FALSE}
cat("\n", strrep("=", 60), "\n")
cat("MODEL TRAINING\n")
cat(strrep("=", 60), "\n")

set.seed(42)
ctrl <- trainControl(
  method = "cv",
  number = 5,
  classProbs = TRUE,
  savePredictions = "final",
  summaryFunction = multiClassSummary,
  verboseIter = FALSE
)

cat("\nNA VALUE CHECK...\n")
na_count <- sum(is.na(train_final))
cat("Total NA count in train_final:", na_count, "\n")

if (na_count > 0) {
  cat("NA values found! Correcting...\n")
  na_columns <- colnames(train_final)[apply(train_final, 2, function(x) any(is.na(x)))]
  cat("Columns containing NA:", paste(na_columns, collapse = ", "), "\n")
  
  numeric_cols <- sapply(train_final, is.numeric)
  for (col in na_columns) {
    if (numeric_cols[col]) {
      train_final[[col]][is.na(train_final[[col]])] <- median(train_final[[col]], na.rm = TRUE)
      cat("  -", col, ": NA values filled with median\n")
    }
  }
  
  if (exists("valid_final")) {
    for (col in na_columns) {
      if (col %in% colnames(valid_final)) {
        if (numeric_cols[col]) {
          valid_final[[col]][is.na(valid_final[[col]])] <- median(valid_final[[col]], na.rm = TRUE)
        }
      }
    }
    cat("NA corrections also applied to valid_final\n")
  }
  
  na_count_after <- sum(is.na(train_final))
  cat("Total NA in train_final after correction:", na_count_after, "\n")
}

cat("\nRANDOM FOREST TRAINING...\n")
set.seed(42)
tryCatch({
  rf_model <- train(
    Status ~ .,
    data = train_final,
    method = "rf",
    trControl = ctrl,
    metric = "Accuracy",
    tuneLength = 3,
    importance = TRUE,
    ntree = 200
  )
  cat("Random Forest results:\n")
  print(rf_model)
  
  if (exists("valid_final") && sum(is.na(valid_final)) == 0) {
    rf_pred <- predict(rf_model, newdata = valid_final)
    cm_rf <- confusionMatrix(rf_pred, valid_final$Status)
    print(cm_rf)
  } else {
    cat("Valid_final contains NA values or doesn't exist\n")
  }
}, error = function(e) {
  cat("Random Forest error:", e$message, "\n")
})

cat("\nLOGISTIC REGRESSION TRAINING...\n")
set.seed(42)
tryCatch({
  if (sum(is.na(train_final)) == 0) {
    log_model <- multinom(
      Status ~ .,
      data = train_final,
      MaxNWts = 10000,
      trace = FALSE
    )
    log_pred <- predict(log_model, newdata = valid_final)
    cm_log <- confusionMatrix(log_pred, valid_final$Status)
    print(cm_log)
  } else {
    cat("Train_final contains NA values, skipping Logistic Regression\n")
    log_model <- NULL
  }
}, error = function(e) {
  cat("Logistic Regression error:", e$message, "\n")
  log_model <- NULL
})

cat("\nGBM TRAINING...\n")
set.seed(42)
ctrl_gbm <- trainControl(
  method = "cv",
  number = 3,
  classProbs = TRUE,
  savePredictions = "final",
  summaryFunction = multiClassSummary,
  verboseIter = FALSE
)

tryCatch({
  if (sum(is.na(train_final)) == 0) {
    gbm_model <- train(
      Status ~ .,
      data = train_final,
      method = "gbm",
      trControl = ctrl_gbm,
      metric = "Accuracy",
      tuneGrid = expand.grid(
        n.trees = c(100, 150),
        interaction.depth = c(3, 5),
        shrinkage = c(0.05, 0.1),
        n.minobsinnode = 10
      ),
      verbose = FALSE
    )
    gbm_pred <- predict(gbm_model, newdata = valid_final)
    cm_gbm <- confusionMatrix(gbm_pred, valid_final$Status)
    print(cm_gbm)
  } else {
    cat("Train_final contains NA values, skipping GBM\n")
    gbm_model <- NULL
  }
}, error = function(e) {
  cat("GBM error:", e$message, "\n")
  gbm_model <- NULL
})
```
## 15) Performance Comparison
- Inputs: cm_rf, cm_log, cm_gbm created in section 14.
- Table: Accuracy and Kappa per model with NA-safe extraction.
- Best model: Prints winner by Accuracy; saves comparison plot.
- Robustness: Filters NA values in plotting.

```{r , message=FALSE, warning=FALSE}
## =========================================================
## 15) Performance Comparison
## =========================================================
cat("\n", strrep("=", 60), "\n")
cat("MODEL PERFORMANCE COMPARISON\n")
cat(strrep("=", 60), "\n")

performance_comparison <- data.frame(
  Model = c("Random Forest", "Logistic Regression", "GBM"),
  Accuracy = NA_real_,
  Kappa = NA_real_
)

# Random Forest
if (exists("cm_rf") && !is.null(cm_rf$overall)) {
  performance_comparison$Accuracy[1] <- ifelse("Accuracy" %in% names(cm_rf$overall),
                                               cm_rf$overall["Accuracy"], NA)
  performance_comparison$Kappa[1] <- ifelse("Kappa" %in% names(cm_rf$overall),
                                            cm_rf$overall["Kappa"], NA)
}

# Logistic Regression
if (exists("cm_log") && !is.null(cm_log$overall)) {
  performance_comparison$Accuracy[2] <- ifelse("Accuracy" %in% names(cm_log$overall),
                                               cm_log$overall["Accuracy"], NA)
  performance_comparison$Kappa[2] <- ifelse("Kappa" %in% names(cm_log$overall),
                                            cm_log$overall["Kappa"], NA)
}

# GBM
if (exists("cm_gbm") && !is.null(cm_gbm$overall)) {
  performance_comparison$Accuracy[3] <- ifelse("Accuracy" %in% names(cm_gbm$overall),
                                               cm_gbm$overall["Accuracy"], NA)
  performance_comparison$Kappa[3] <- ifelse("Kappa" %in% names(cm_gbm$overall),
                                            cm_gbm$overall["Kappa"], NA)
}

performance_comparison$Accuracy <- round(performance_comparison$Accuracy, 4)
performance_comparison$Kappa <- round(performance_comparison$Kappa, 4)

cat("\nPERFORMANCE TABLE:\n")
print(performance_comparison)

valid_rows <- !is.na(performance_comparison$Accuracy)
if (any(valid_rows)) {
  best_model_idx <- which.max(performance_comparison$Accuracy[valid_rows])
  best_model_name <- performance_comparison$Model[valid_rows][best_model_idx]
  cat("\nBEST MODEL:", best_model_name,
      "\n   Accuracy:", performance_comparison$Accuracy[valid_rows][best_model_idx],
      "\n   Kappa:", performance_comparison$Kappa[valid_rows][best_model_idx], "\n")
} else {
  cat("\nPerformance could not be calculated for any models\n")
}

performance_long <- performance_comparison %>%
  filter(!is.na(Accuracy) | !is.na(Kappa)) %>%
  tidyr::pivot_longer(cols = c(Accuracy, Kappa), names_to = "Metric", values_to = "Value") %>%
  dplyr::filter(!is.na(Value))

if (nrow(performance_long) > 0) {
  p_perf <- ggplot(performance_long, aes(x = Model, y = Value, fill = Metric)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    geom_text(aes(label = sprintf("%.3f", Value)),
              position = position_dodge(width = 0.9), vjust = -0.5, size = 3.5) +
    scale_fill_brewer(palette = "Set2") +
    labs(
      title = "Model Performance Comparison",
      subtitle = "After Correlation Analysis Based Feature Engineering",
      x = "Model",
      y = "Value",
      fill = "Metric"
    ) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"),
          plot.subtitle = element_text(hjust = 0.5))
  
  print(p_perf)
  ggsave(file.path(PROJECT_DIR, "model_performance_comparison_final.png"),
         plot = p_perf, width = 12, height = 8)
}
```
## 16) Feature Importance
- Model: rf_model from section 14.
- Output: Top-20 features bar chart and CSVs (all and top-20).
- Robustness: Handles multi-class varImp, adds short names for readability.

```{r , message=FALSE, warning=FALSE}
## =========================================================
## 16) Feature Importance Analysis
## =========================================================
cat("\n", strrep("=", 60), "\n")
cat("FEATURE IMPORTANCE ANALYSIS\n")
cat(strrep("=", 60), "\n")

if (exists("rf_model") && "finalModel" %in% names(rf_model)) {
  tryCatch({
    rf_importance <- varImp(rf_model, scale = TRUE)
    cat("\nRANDOM FOREST - TOP 20 IMPORTANT VARIABLES:\n")
    
    imp_df <- as.data.frame(rf_importance$importance)
    if (!"Feature" %in% colnames(imp_df)) imp_df$Feature <- rownames(imp_df)
    
    if (ncol(imp_df) > 2) {
      imp_values <- imp_df[, 1:(ncol(imp_df)-1)]
      imp_df$Overall <- rowMeans(imp_values, na.rm = TRUE)
    } else if (ncol(imp_df) == 2) {
      names(imp_df) <- c("Overall", "Feature")
    }
    
    imp_df <- imp_df[order(-imp_df$Overall), ]
    imp_df$Short_Name <- sapply(imp_df$Feature, function(x) {
      if (nchar(x) > 40) paste0(substr(x, 1, 37), "...") else x
    })
    
    top_20 <- head(imp_df, 20)
    cat("\nTop 20 Feature Importance:\n")
    print(top_20[, c("Feature", "Overall")])
    
    if (nrow(top_20) > 0) {
      p_imp <- ggplot(top_20, aes(x = reorder(Short_Name, Overall), y = Overall)) +
        geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
        coord_flip() +
        labs(
          title = "Top 20 Feature Importance (Random Forest)",
          subtitle = "Correlation Analysis Based Model",
          x = "Feature",
          y = "Importance Score"
        ) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, face = "bold"),
              plot.subtitle = element_text(hjust = 0.5))
      
      print(p_imp)
      ggsave(file.path(PROJECT_DIR, "feature_importance_top20.png"),
             plot = p_imp, width = 12, height = 8)
      
      write.csv(imp_df, file.path(PROJECT_DIR, "feature_importance_all.csv"), row.names = FALSE)
      write.csv(top_20, file.path(PROJECT_DIR, "feature_importance_top20.csv"), row.names = FALSE)
      cat("\nFeature importance saved.\n")
    }
  }, error = function(e) {
    cat("Feature importance could not be calculated:", e$message, "\n")
  })
} else {
  cat("rf_model not available; skipping feature importance.\n")
}
```
## 17) Save Objects (For Test Phase)
- **Purpose:** Persist all critical artifacts produced during the training phase so they can be reused consistently during the test phase.
- **Saved Objects:**
  - **Models:** Random Forest, Logistic Regression, GBM
  - **Preprocessing:** Dummy encoder object, categorical column list
  - **Datasets:** Training and modeling master tables
  - **Correlation Analysis:** Raw, post–feature engineering, and cleaned correlation outputs
  - **Performance Metrics:** Model performance comparison table
- **Output File:**
  - `train_models_final.RData`
- **Benefit:**
  - Ensures a **leak-free test pipeline** by reusing the exact same preprocessing and trained models.

```{r , message=FALSE, warning=FALSE}
cat("\n", strrep("=", 60), "\n")
cat("SAVING OBJECTS...\n")
cat(strrep("=", 60), "\n")

save(
  # Model objects
  train_final,
  dummies,
  cat_cols,
  rf_model,
  log_model,
  gbm_model,
  
  # Original data
  master_train,
  master_model,
  
  # Correlation analysis results
  cor_matrix_raw,
  cor_matrix_final,
  cor_matrix_clean,
  high_corr_df_raw,
  high_corr_df_final,
  high_corr_df_clean,
  status_correlations_raw,
  status_correlations_final,
  status_correlations_clean,
  
  # Performance results
  performance_comparison,
  
  file = file.path(PROJECT_DIR, "train_models_final.RData")
)

cat("\nCREATED FILES:\n")
cat("1. train_models_final.RData - Main model objects\n")
cat("2. correlation_raw_before_fe.png - Raw correlation heatmap\n")
cat("3. correlation_final_after_fe.png - Post-feature engineering heatmap\n")
cat("4. correlation_clean_after_removal.png - Post-cleaning heatmap\n")
cat("5. high_correlation_raw.csv - Raw high correlations\n")
cat("6. high_correlation_final.csv - Post-feature engineering high correlations\n")
cat("7. high_correlation_clean.csv - Post-cleaning high correlations\n")
cat("8. status_correlation.png - Status correlation graph\n")
cat("9. status_correlation_clean.png - Post-cleaning status correlation\n")
cat("10. new_features_correlation.png - New features correlation\n")
cat("11. model_performance_comparison_final.png - Model performance\n")
cat("12. feature_importance_top20.png - Feature importance\n")
cat("13. feature_importance_all.csv - All feature importance\n")
cat("14. feature_importance_top20.csv - Top 20 feature importance\n")
```

## 18) Summary Report
- **Purpose:** Generate a concise, human-readable summary of the entire training process directly in the console.
- **Includes:**
  - Dataset sizes (train / validation)
  - Number of features before and after feature engineering and cleaning
  - High-correlation reduction statistics and success rate
  - List of newly engineered features
  - Model-level performance metrics (Accuracy and Kappa)
- **Benefit:**
  - Provides a **single-glance validation** of the training pipeline.
  - Ready-to-use summary for reporting and presentations.

```{r , message=FALSE, warning=FALSE}
cat("\n", strrep("=", 60), "\n")
cat("SUMMARY REPORT\n")
cat(strrep("=", 60), "\n")

cat("\nDATA SUMMARY:\n")
cat("Total Customers:", nrow(master_train), "\n")
cat("Train Set:", nrow(train_data), "\n")
cat("Validation Set:", nrow(valid_data), "\n")
cat("Total Features:", ncol(train_final) - 1, "\n")

cat("\nCORRELATION ANALYSIS COMPARISON:\n")
cat("1. Before Feature Engineering:", ncol(numeric_vars_raw), "variables\n")
cat("2. After Feature Engineering:", ncol(numeric_vars_final), "variables\n")
cat("3. After Cleaning:", ncol(numeric_vars_clean), "variables\n")

if (exists("high_corr_df_raw") && exists("high_corr_df_clean")) {
  cat("\nHIGH CORRELATION REDUCTION SUCCESS:\n")
  cat("   Initial:", nrow(high_corr_df_raw), "high correlation pairs\n")
  cat("   After Cleaning:", nrow(high_corr_df_clean), "high correlation pairs\n")
  
  if (nrow(high_corr_df_raw) > 0 && nrow(high_corr_df_clean) > 0) {
    reduction <- round(100 * (1 - nrow(high_corr_df_clean) / nrow(high_corr_df_raw)), 1)
    cat("   Reduction rate: %", reduction, "\n")
  }
}

cat("\nNEW FEATURES:\n")
cat("1. Satisfaction_Composite - Satisfaction composite score\n")
cat("2. Support_Burden - Support burden\n")
cat("3. Revenue_per_Month - Monthly revenue efficiency\n")
cat("4. Bug_Frequency - Bug frequency\n")
cat("5. Engagement_Ratio - Engagement ratio\n")
cat("6. MRR_Ratio - MRR ratio\n")

cat("\nMODEL RESULTS:\n")
for (i in 1:nrow(performance_comparison)) {
  if (!is.na(performance_comparison$Accuracy[i])) {
    cat(sprintf("%-20s: Accuracy = %.3f, Kappa = %.3f\n", 
                performance_comparison$Model[i],
                performance_comparison$Accuracy[i],
                performance_comparison$Kappa[i]))
  }
}

cat("\n", strrep("=", 60), "\n")
cat("TRAIN PHASE SUCCESSFULLY COMPLETED!\n")
cat(strrep("=", 60), "\n")

cat("\nCompletion Time:", Sys.time(), "\n")
cat("Project Directory:", PROJECT_DIR, "\n")
cat("Ready to proceed to test phase!\n")
```

## 19) Advanced Train Visualizations (Different Techniques)
- **Purpose:** Go beyond standard plots to improve **interpretability and segment-level insights**.
- **Outputs:**
  1. **Lollipop Chart**
     - Visualizes the top predictors with the strongest correlation to customer status.
     - Highlights both magnitude and direction of influence.
  2. **Radar (Spider) Chart**
     - Compares normalized behavioral and financial profiles across customer status groups.
     - Useful for understanding multi-dimensional differences between segments.
- **Saved Files:**
  - `train_lollipop_influence.png`
  - `train_status_radar.png`
- **Benefit:**
  - Enhances **explainability** and stakeholder-friendly storytelling.

```{r , message=FALSE, warning=FALSE}
cat("\n", strrep("=", 60), "\n")
cat("GENERATING ADVANCED DIAGNOSTIC PLOTS\n")
cat(strrep("=", 60), "\n")

# 19.1 Lollipop Chart: Top 15 Variables by Status Correlation
if (exists("status_correlations_clean")) {
  p_lollipop <- status_correlations_clean %>%
    head(15) %>%
    ggplot(aes(x = reorder(Variable, Abs_Correlation), y = Correlation_with_Status)) +
    geom_segment(aes(xend = Variable, yend = 0), color = "grey") +
    geom_point(aes(color = Correlation_with_Status > 0), size = 4) +
    scale_color_manual(values = c("TRUE" = "#2ecc71", "FALSE" = "#e74c3c"), 
                       labels = c("Negative", "Positive"), name = "Correlation Dir.") +
    coord_flip() +
    labs(title = "Feature Influence on Customer Status",
         subtitle = "Lollipop Chart: Strongest Predictors (Clean Data)",
         x = "Features", y = "Correlation Coefficient") +
    theme_minimal()
  
  print(p_lollipop)
  ggsave(file.path(PROJECT_DIR, "train_lollipop_influence.png"), p_lollipop, width = 10, height = 7)
}

# 19.2 Radar Chart (Spider Plot) - Status Characterization
cat("\nGenerating Status Profile Radar Analysis...\n")

radar_data <- master_train %>%
  group_by(Status) %>%
  summarise(
    Avg_Satisfaction = mean(Satisfaction_Composite, na.rm = TRUE),
    Avg_Bug_Freq = mean(Bug_Frequency, na.rm = TRUE),
    Avg_Engagement = mean(Engagement_Ratio, na.rm = TRUE),
    Avg_Revenue = mean(Revenue_per_Month, na.rm = TRUE),
    Avg_MRR_Ratio = mean(MRR_Ratio, na.rm = TRUE)
  ) %>%
  mutate(across(-Status, rescale)) # Normalize to 0-1 range

radar_long <- radar_data %>%
  pivot_longer(cols = -Status, names_to = "Metric", values_to = "Value")

p_radar <- ggplot(radar_long, aes(x = Metric, y = Value, group = Status, color = Status)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  coord_polar() +
  labs(title = "Customer Status Profiles",
       subtitle = "Comparing normalized metrics across groups") +
  theme_light() +
  theme(axis.text.y = element_blank())

print(p_radar)
ggsave(file.path(PROJECT_DIR, "train_status_radar.png"), p_radar, width = 8, height = 8)

```

## 20) Multi-Model Confusion Matrix Visualization
- **Purpose:** Provide a detailed, visual comparison of classification performance across all trained models.
- **Process:**
  - Builds confusion matrices for Random Forest, Logistic Regression, and GBM.
  - Displays both absolute counts and row-wise percentages.
- **Outputs:**
  - Individual confusion matrix heatmaps per model.
- **Saved Files:**
  - `confusion_matrix_random_forest.png`
  - `confusion_matrix_logistic_regression.png`
  - `confusion_matrix_gbm.png`
- **Benefit:**
  - Makes **error patterns and class-level weaknesses** immediately visible.
  - Supports informed model selection beyond accuracy alone.

```{r , message=FALSE, warning=FALSE}
cat("\n", strrep("=", 60), "\n")
cat("GENERATING CONFUSION MATRICES FOR ALL MODELS\n")
cat(strrep("=", 60), "\n")

# Prepare model and CM list
cm_list <- list()
if (exists("cm_rf"))  cm_list[["Random_Forest"]] <- cm_rf
if (exists("cm_log")) cm_list[["Logistic_Regression"]] <- cm_log
if (exists("cm_gbm")) cm_list[["GBM"]] <- cm_gbm

if (length(cm_list) > 0) {
  for (model_name in names(cm_list)) {
    
    # 1. Prepare Data
    cm_obj <- cm_list[[model_name]]
    cm_df <- as.data.frame(cm_obj$table)
    
    # Calculate percentage values (for better analysis)
    cm_df <- cm_df %>%
      group_by(Reference) %>%
      mutate(Pct = Freq / sum(Freq) * 100) %>%
      ungroup()
    
    # 2. Visualization
    p_cm_model <- ggplot(cm_df, aes(x = Reference, y = Prediction, fill = Freq)) +
      geom_tile(color = "white", size = 1) +
      geom_text(aes(label = sprintf("%d\n(%.1f%%)", Freq, Pct)), 
                color = "white", fontface = "bold", size = 5) +
      scale_fill_gradient(low = "#2c3e50", high = "#2980b9") +
      labs(
        title = paste("Confusion Matrix:", gsub("_", " ", model_name)),
        subtitle = "Reference: Actual Values | Prediction: Model Guesses",
        x = "Actual Status (Reference)",
        y = "Predicted Status (Prediction)",
        fill = "Count"
      ) +
      theme_minimal() +
      theme(
        plot.title = element_text(face = "bold", size = 14),
        axis.text = element_text(size = 11, face = "bold"),
        panel.grid = element_blank()
      )
    
    # 3. Print and Save
    print(p_cm_model)
    filename <- paste0("confusion_matrix_", tolower(model_name), ".png")
    ggsave(file.path(PROJECT_DIR, filename), p_cm_model, width = 8, height = 6)
    
    cat("Saved:", filename, "\n")
  }
} else {
  cat("No Confusion Matrix objects found (cm_rf, cm_log, or cm_gbm).\n")
}
```

## 21) Multi-Class ROC & AUC Analysis
- **Purpose:** Evaluate model discrimination power using **One-vs-Rest ROC and AUC metrics** for multi-class classification.
- **Process:**
  - Computes ROC curves separately for each class and model.
  - Aggregates AUC scores for comparative analysis.
- **Outputs:**
  - Faceted ROC curves by class
  - Bar charts summarizing AUC scores per model and class
- **Saved Files:**
  - `train_multi_roc_curves.png`
  - `train_auc_comparison.png`
- **Benefit:**
  - Provides a **threshold-independent performance assessment**.
  - Highlights which models perform best for specific customer segments.

```{r , message=FALSE, warning=FALSE}
cat("\n", strrep("=", 60), "\n")
cat("GENERATING ROC CURVES AND AUC SCORES\n")
cat(strrep("=", 60), "\n")

# Get prediction probabilities
prob_list <- list()
if (exists("rf_model"))  prob_list[["Random_Forest"]] <- predict(rf_model, newdata = valid_final, type = "prob")
if (exists("log_model")) prob_list[["Logistic_Reg"]] <- predict(log_model, newdata = valid_final, type = "probs")
if (exists("gbm_model")) prob_list[["GBM"]] <- predict(gbm_model, newdata = valid_final, type = "prob")

# List to store ROC data
roc_plot_data <- data.frame()

if (length(prob_list) > 0) {
  for (model_name in names(prob_list)) {
    probs <- prob_list[[model_name]]
    
    # Calculate ROC for each class (One-vs-Rest)
    classes <- levels(valid_final$Status)
    
    for (cl in classes) {
      # Convert actual values to binary (Selected class vs Others)
      binary_actual <- ifelse(valid_final$Status == cl, 1, 0)
      current_prob <- probs[, cl]
      
      roc_obj <- roc(binary_actual, current_prob, quiet = TRUE)
      
      # Add to plot data
      tmp_df <- data.frame(
        Specificity = roc_obj$specificities,
        Sensitivity = roc_obj$sensitivities,
        Model = model_name,
        Class = cl,
        AUC = as.numeric(roc_obj$auc)
      )
      roc_plot_data <- rbind(roc_plot_data, tmp_df)
    }
  }
  
  # 1. Separate ROC for Each Class (Facet Wrap)
  p_roc <- ggplot(roc_plot_data, aes(x = 1 - Specificity, y = Sensitivity, color = Model)) +
    geom_line(size = 1.2) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey") +
    facet_wrap(~Class) +
    labs(
      title = "Multi-class ROC Curves (One-vs-Rest)",
      subtitle = "Performance comparison across different customer segments",
      x = "False Positive Rate (1 - Specificity)",
      y = "True Positive Rate (Sensitivity)"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))
  
  print(p_roc)
  ggsave(file.path(PROJECT_DIR, "train_multi_roc_curves.png"), p_roc, width = 12, height = 7)
  
  # 2. AUC Score Summary Table Visualization
  auc_summary <- roc_plot_data %>%
    group_by(Model, Class) %>%
    summarise(AUC_Score = max(AUC), .groups = "drop")
  
  p_auc_bar <- ggplot(auc_summary, aes(x = Class, y = AUC_Score, fill = Model)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
    geom_text(aes(label = round(AUC_Score, 3)), 
              position = position_dodge(width = 0.8), vjust = -0.5, size = 3) +
    ylim(0, 1.1) +
    labs(
      title = "AUC Scores per Class and Model",
      subtitle = "Values closer to 1.0 indicate better separation power",
      x = "Customer Status Class",
      y = "AUC Score"
    ) +
    theme_minimal()
  
  print(p_auc_bar)
  ggsave(file.path(PROJECT_DIR, "train_auc_comparison.png"), p_auc_bar, width = 10, height = 6)
  
  cat("ROC and AUC analysis completed. Check your folder for .png files.\n")
} else {
  cat("Probabilities could not be retrieved for ROC analysis.\n")
}
cat("\nAdvanced visualizations saved to your project directory.\n")

```

## 22) All Models Feature Importance Comparison
- **Purpose:** Compare feature importance patterns across different modeling techniques.
- **Methodology:**
  - **Random Forest:** Mean decrease–based importance
  - **GBM:** Model-derived relative importance
  - **Logistic Regression:** Absolute coefficient values (normalized) as an importance proxy
- **Visualization:**
  - Top 15 features per model displayed in faceted bar charts.
- **Saved File:**
  - `all_models_feature_importance.png`
- **Benefit:**
  - Reveals **model-specific drivers** and consensus predictors.
  - Supports interpretability and business-level feature validation.

```{r , message=FALSE, warning=FALSE}
cat("\n", strrep("=", 60), "\n")
cat("GENERATING ALL-IN-ONE FEATURE IMPORTANCE COMPARISON\n")
cat(strrep("=", 60), "\n")

all_imp_data <- data.frame()

# 1. Random Forest Importance
if (exists("rf_model")) {
  rf_imp <- varImp(rf_model, scale = TRUE)$importance
  rf_df <- data.frame(
    Feature = rownames(rf_imp),
    Importance = rowMeans(rf_imp),
    Model = "Random Forest"
  )
  all_imp_data <- rbind(all_imp_data, rf_df)
}

# 2. GBM Importance
if (exists("gbm_model")) {
  gbm_imp <- varImp(gbm_model, scale = TRUE)$importance
  gbm_df <- data.frame(
    Feature = rownames(gbm_imp),
    Importance = gbm_imp$Overall,
    Model = "GBM"
  )
  all_imp_data <- rbind(all_imp_data, gbm_df)
}

# 3. Logistic Regression (Absolute Coefficients as Proxy for Importance)
if (exists("log_model") && !is.null(log_model)) {
  # In logistic regression, absolute coefficient values represent importance
  log_coefs <- abs(summary(log_model)$coefficients)
  # If multiple classes, take average of coefficients
  if (is.matrix(log_coefs)) {
    log_imp_vals <- colMeans(log_coefs)
  } else {
    log_imp_vals <- log_coefs
  }
  
  log_df <- data.frame(
    Feature = names(log_imp_vals),
    Importance = log_imp_vals,
    Model = "Logistic Regression"
  )
  # Normalize to 0-100 scale
  log_df$Importance <- (log_df$Importance / max(log_df$Importance)) * 100
  all_imp_data <- rbind(all_imp_data, log_df %>% filter(Feature != "(Intercept)"))
}

# 4. Visualization
if (nrow(all_imp_data) > 0) {
  # Select top 15 features for each model
  top_all_imp <- all_imp_data %>%
    group_by(Model) %>%
    arrange(desc(Importance)) %>%
    slice_head(n = 15) %>%
    ungroup()
  
  p_all_imp <- ggplot(top_all_imp, aes(x = reorder_within(Feature, Importance, Model), 
                                       y = Importance, fill = Importance)) +
    geom_bar(stat = "identity") +
    scale_x_reordered() +
    scale_fill_gradient(low = "#95a5a6", high = "#2c3e50") +
    coord_flip() +
    facet_wrap(~Model, scales = "free") +
    labs(
      title = "Feature Importance Comparison: All Models",
      subtitle = "Top 15 features ranked by relative impact per model",
      x = "Predictor Variables",
      y = "Normalized Importance (0-100)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 16),
      strip.text = element_text(face = "bold", size = 12),
      legend.position = "none"
    )
  
  print(p_all_imp)
  ggsave(file.path(PROJECT_DIR, "all_models_feature_importance.png"), 
         plot = p_all_imp, width = 16, height = 9)
  
  cat("Combined Feature Importance chart saved: all_models_feature_importance.png\n")
}
```